---
title: "ESM 244 LAb 2 DN"
author: "Derek Nguyen"
date: "1/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(naniar)
library(VIM)
library(skimr)
library(ggfortify)
library(here)
```

## 1. Get some data

```{r}
ca_pb <- read_csv(here("data", "ca_pollution_burden.csv"))

ca_dem <- read_csv(here("data", "ca_census_demographics_2010.csv"))
```

## 2. Some data cleaning and wrangling

```{r}
ca_pb_nopct <- ca_pb %>% 
  clean_names() %>% 
  select(-contains("perc")) %>% 
  select(-contains("pctl")) %>% 
  select(-longitude, - latitude)
```

```{r}
ca_dem_clean <- ca_dem %>% 
  clean_names()
```

## 3. Pollution burdens indicators only

```{r}
ca_pb_subset <- ca_pb_nopct %>% 
  select(ozone:solid_waste, asthma:housing_burden)
```



Now, lets do PCA
use the `prcomp()` function
```{r, eval= FALSE}

# this gives an error because there are missing values!!
pb_pca <- prcomp(ca_pb_subset, scale = TRUE)


```

Aside, exploring missingness!
```{r}
summary(ca_pb_subset)

gg_miss_var(ca_pb_subset)

matrixplot(ca_pb_subset, sortby = "poverty")
```

Remove (complete case deletion) any observation with ANY NA values:

```{r}
ca_pb_nona <- ca_pb_subset %>% 
  drop_na()

summary(ca_pb_nona)

skim(ca_pb_nona)
```

Now, no NA values, cool set to do PCA (otherwise, consider imputation, etc)

```{r}
my_ca_pca <- prcomp(ca_pb_nona, scale = TRUE)

my_ca_pca
summary(my_ca_pca)

#biplot(my_ca_pca)

# actually lets use autoplot

my_biplot <- autoplot(my_ca_pca,
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) +
  scale_y_continuous(limits = c(-0.05,0.05))

my_biplot
```

## 4. PCA for pollution burden + demographics

Join together the ca_dem_clean with ca_pb_nopct

```{r}
ca_df <- ca_dem_clean %>% 
  inner_join(ca_pb_nopct, by = c("census_tract_number" = "census_tract"))


#now drop NAs by complete case deletion:
ca_nona <- ca_df %>% 
  drop_na()
```

MAke a subset of ca_nona that only includes:
- white_percent
- elderly_65_percent
- pm2_5
- pesticides
- traffic
- asthma
- cardiovascular_disease
- poverty

```{r}
my_sub <- ca_nona %>% 
  select(white_percent, elderly_65_percent, pm2_5, pesticides, traffic, asthma, cardiovascular_disease, poverty)
```



PCA:

```{r}
my_dem_pca <- prcomp(my_sub, scale = TRUE)

my_dem_biplot <- autoplot(my_dem_pca,
                      colour = NA,
                      loadings.label = TRUE,
                      loadings.label.size = 3,
                      loadings.label.colour = "black",
                      loadings.label.repel = TRUE) +
  scale_y_continuous(limits = c(-0.05,0.05)) +
  theme_minimal()

my_dem_biplot
```

















